Seminar: Programming models and Code Generation\\ Winter Term 2014/2015\\ {\bf Quasi deterministic parallel programming using LVars}

### 2015-01-10 draft.tex
new

To describe Lvar formally, a parallel call-by-value $\lambda$-calculus named $\lambdaLVar$ is introduced in \cite{lkuper} which is extended with a $store$ and communication primitives $put$ and $get$ that operate on data in the $store$. The definition of $\lambdaLVar$ is parameterized by the choice of lattice and therefore $\lambdaLVar$ is actually a family of languages, rather than a single language. An LVar may have an arbitrary number of states forming a set $D$, which is partially ordered by a relation $\userleq$. Formally, $D$ is a \emph{bounded-join-semilattice} augmented with a greatest element $\top$. 
\begin{itemize}
\item $D$ has a least element $\bot$, representing the initial “empty” state of an Lvar,
\item $D$ has a greatest element $\top$, representing the “error” state that results from conflicting updates to an Lvar,
\item $D$ comes equipped with a partial order $\userleq$, where $\bot \userleq d \userleq \top$ for all $d \in D$.
\item Every pair of elements in $D$ has a least upper bound (lub) written $\sqcup$ which means it is possible for two sub-computations to independently update an LVar, and then deterministically merge the results by taking the lub of the resulting two states.
\end{itemize}
% free floating figure using width of full page, to be put on [t]op.
\begin{figure}
\centerline{
\includegraphics[width=1.0\columnwidth]{Figures/lattice.jpg}
}
\caption{Example lattices\cite{lkuper}: (a) Positive integers ordered by $\leq$ (b) IVar containing a natural number (c) Pair of natural number-valued IVars}
\label{Figure1}
\end{figure}
Data structures like arrays, tree, maps , streams etc. can be represented  as lattice. $Figure$ $2$ gives three examples of lattices for common data structures.

@@

\textbf{\textit{Proof of determinism:}} The key to the determinism is the frame rule given in \ref{f:frame-rule}. A frame property captures the idea of local reasoning about programs that alter state\cite{lindsey}. Here, $C$ is a program, and $\{p\} C \{q\}$ says that if the assertion $p$ is true before $C$ runs, then the assertion $q$ will be true afterwards. For example, $p$ and $q$  might describe the state of the heap before and after it is updated by $C$. It tells that running $C$ starting from a state satisfying the assertion $p$ $\ast$ $r$ will result in a state satisfying the assertion $q$ $\ast$ $r$. The assertion $p$ $\ast$ $r$ is satisfied by a heap if the heap can be split into two non-overlapping parts satisfying $p$ and $r$, respectively.

@@

\textbf{\textit{Key Ideas:}} \textit{Atoms and Idempotence:} \textit{Atoms} are elements not equal to $\bot$ but whose only smaller element is $\bot$. Lattices in which every element is the LUB of some set of atoms are called atomistic, and most application-specific lattices used by LVish programs have this property. \ref{Figure1} is an example for an atomistic lattice where the atoms are singleton elements and all the elements can be obtained as LUB of these atoms. Data structures corresponding to such type of lattices usually expose operations that work at the atom level, semantically limiting $puts$ to atoms, $gets$ to threshold sets of atoms, and event sets to sets of atoms. LVish facilitates good performance for atomistic lattices by associating LVars with a set of deltas(changes), as well a lattice. For atomistic lattices, the deltas are just the atoms, for a set, a delta is an element; for a map, a key/value pair. Deltas allow easy and efficient communication of changes between $puts$ and $gets$/handlers. \par

@@

\textbf{\textit{Evaluation: k-CFA study}} In \cite{lkuper2} a case study on k-CFA (parallelizing control flow analysis) is done to evaluate the performance of the Lvish implementation. K-CFA provides a hierarchy of increasingly precise methods to compute the control flow of a program or precisely the flow of values to expressions. \ref{Figure2} shows performance data for LVish implementation using lock-free data structures on 2 benchmarks. One is the blur benchmark based on k=2. Another is the synthetic benchmark notChain which is simply a chain of 300 not functions that negates the benefit of the sharing approach in the LVish model.\\ \\

>>>>

To describe Lvar formally, a parallel call-by-value $\lambda$-calculus named $\lambdaLVar$ is introduced in \cite{lkuper} which is extended with a $store$ and communication primitives $put$ and $get$ that operate on data in the $store$. The definition of $\lambdaLVar$ is parameterized by the choice of lattice and therefore $\lambdaLVar$ is actually a family of languages, rather than a single language. An LVar may have an arbitrary number of states forming a set $D$, which is partially ordered by a relation $\userleq$. Formally, $D$ is a \emph{bounded-join-semilattice} augmented with a greatest element $\top$. 
\begin{itemize}
\item $D$ has a least element $\bot$, representing the initial “empty” state of an Lvar,
\item $D$ has a greatest element $\top$, representing the “error” state that results from conflicting updates to an Lvar,
\item $D$ comes equipped with a partial order $\userleq$, where $\bot \userleq d \userleq \top$ for all $d \in D$.
\item Every pair of elements in $D$ has a least upper bound (lub) written $\sqcup$ which means it is possible for two sub-computations to independently update an LVar, and then deterministically merge the results by taking the lub of the resulting two states.
\end{itemize}
% free floating figure using width of full page, to be put on [t]op.
\begin{figure}
\centerline{
\includegraphics[width=1.0\columnwidth]{Figures/lattice.jpg}
}
\caption{Example lattices\cite{lkuper}: (a) Positive integers ordered by $\leq$ (b) IVar containing a natural number (c) Pair of natural number-valued IVars}
\label{Figure1}
\end{figure}
Data structures like arrays, tree, maps , streams etc. can be represented  as lattice. $Figure\ 1$ \ref{Figure1} gives three examples of lattices for common data structures.

@@

\textbf{\textit{Proof of determinism:}} The key to the determinism is the frame rule given in $Figure\ 2$ \ref{f:frame-rule}. A frame property captures the idea of local reasoning about programs that alter state\cite{lindsey}. Here, $C$ is a program, and $\{p\} C \{q\}$ says that if the assertion $p$ is true before $C$ runs, then the assertion $q$ will be true afterwards. For example, $p$ and $q$  might describe the state of the heap before and after it is updated by $C$. It tells that running $C$ starting from a state satisfying the assertion $p$ $\ast$ $r$ will result in a state satisfying the assertion $q$ $\ast$ $r$. The assertion $p$ $\ast$ $r$ is satisfied by a heap if the heap can be split into two non-overlapping parts satisfying $p$ and $r$, respectively.

@@

\textbf{\textit{Key Ideas:}} \textit{Atoms and Idempotence:} \textit{Atoms} are elements not equal to $\bot$ but whose only smaller element is $\bot$. Lattices in which every element is the LUB of some set of atoms are called atomistic, and most application-specific lattices used by LVish programs have this property. $Figure\ 1$ \ref{Figure1} is an example for an atomistic lattice where the atoms are singleton elements and all the elements can be obtained as LUB of these atoms. Data structures corresponding to such type of lattices usually expose operations that work at the atom level, semantically limiting $puts$ to atoms, $gets$ to threshold sets of atoms, and event sets to sets of atoms. LVish facilitates good performance for atomistic lattices by associating LVars with a set of deltas(changes), as well a lattice. For atomistic lattices, the deltas are just the atoms, for a set, a delta is an element; for a map, a key/value pair. Deltas allow easy and efficient communication of changes between $puts$ and $gets$/handlers. \par

@@

\textbf{\textit{Evaluation: k-CFA study}} In \cite{lkuper2} a case study on k-CFA (parallelizing control flow analysis) is done to evaluate the performance of the Lvish implementation. K-CFA provides a hierarchy of increasingly precise methods to compute the control flow of a program or precisely the flow of values to expressions. $Figure\ 3$ \ref{Figure2} shows performance data for LVish implementation using lock-free data structures on 2 benchmarks. One is the blur benchmark based on k=2. Another is the synthetic benchmark notChain which is simply a chain of 300 not functions that negates the benefit of the sharing approach in the LVish model.\\ \\
